---
title: "Summarizing The Weather"
date: "4/23/2021"
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = paste0(
        xfun::sans_ext(input), '.pdf'
      ),
      output_dir = "../../docs"
    )
  })
output: 
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 2
    fig_width: 7
    fig_height: 6
    fig_caption: true
    df_print: kable
    highlight: tango
    keep_tex: true
fontsize: 11pt
geometry: margin=1in
---

# Setup and Libraries

```{r}
library(magrittr)
library(lubridate)
```

# Introduction
This Code Clinic problem is about calculating statistics from a data set.
It's easy stuff, but presents a good example of how different languages
accomplish common tasks.


# Import the source data
The data set is weather data captured from Lake Pend O'Reille
in Northern Idaho. We have almost 20 megabytes of data from the
years 2012 thorugh 2015. That data is available in the folder with
other exercise files. Each observation in the data includes several
variables and the data is straightforward.

```{r}
mytempfile <- tempfile()

readOneFile <- function(dataPath) {
  read.table(dataPath,
             header = TRUE,
             stringsAsFactors = FALSE)
}

myProgressBar <- txtProgressBar(min = 2012, max = 2015, style = 3)

for (dataYear in 2012:2015) {
  
  dataPath <-
    paste0(
      "https://raw.githubusercontent.com/lyndadotcom/LPO_weatherdata/master/Environmental_Data_Deep_Moor_",
      dataYear,
      ".txt"
    )
  
  if (exists("LPO_weather_data")) {
    mytempfile <- readOneFile(dataPath)
    LPO_weather_data <- rbind(LPO_weather_data, mytempfile)
  } else {
    LPO_weather_data <- readOneFile(dataPath)
  }
  setTxtProgressBar(myProgressBar, value = dataYear)
  
}

# confirm the results of the import
head(LPO_weather_data, n = 3)
tail(LPO_weather_data, n = 3)

print(paste("Number of rows imported: ", nrow(LPO_weather_data)))
```

# Calculate the Coefficient of Barometric Pressure 
The problem is simple: Write a function that accepts ...
a beginning date and time
...and...
an ending date and time...

```{r}
startDateTime <- "2014-01-02 12:03:34"
endDateTime <- "2014-01-04 12:03:34"
```

...then...
inclusive of those dates and times return the coefficient of the
slope of barometric pressure.

helper function to get a subset of LPO_weather_data
observations are the date range
variables are barometric pressure, date, and time

```{r}
getBaromPressures <- function(dateTimeInterval) {
  subset(
    LPO_weather_data,
    ymd_hms(paste(date, time)) %within% dateTimeInterval,
    select = c(Barometric_Press, date, time)
  )
}

calculateBaroPress <- function(startDateTime, endDateTime) {
  dateTimeInterval <- interval(ymd_hms(startDateTime),
                               ymd_hms(endDateTime))
  
  baroPress <- getBaromPressures(dateTimeInterval)
  
  slope <- ymd_hms(paste(baroPress$date, baroPress$time))
  
  lm(Barometric_Press ~ slope, data = baroPress)

}

calculateBaroPress(startDateTime, endDateTime)
```

A rising slope indicates an increasing barometric pressure,
which typically means fair and sunny weather. A falling slope
indicates a decreasing barometric pressure, which typically means
stormy weather.

We're only asking for the coefficient – but some may choose
to generate a graph of the results as well.

# Graph Barometric Pressure 

```{r}
graphBaroPressure <- function(startDateTime, endDateTime ) {
  
  dateTimeInterval <- interval(ymd_hms(startDateTime),
                               ymd_hms(endDateTime))
  
  baroPress <- getBaromPressures(dateTimeInterval)
  
  thisDateTime <- ymd_hms(paste(baroPress$date, baroPress$time))
  
  plot(
    x = thisDateTime,
    y = baroPress$Barometric_Press,
    xlab = "Date and Time",
    ylab = "Barometric Pressure",
    main = paste(
      "Barometric Pressure from ",
      ymd_hms(startDateTime),
      "to",
      ymd_hms(endDateTime)
    )
  )
  abline(calculateBaroPress(startDateTime, endDateTime), col = "red")
}

graphBaroPressure(startDateTime, endDateTime)
```

